Analyzing Performance Bottleneck in Graph Neural Network Training: An Experimental View

### 2020-07-15 Manuscript/sec4-experiments.tex
update初稿

The goal of this experiment is to verify the accuracy of complexity analysis in Table~\ref{tab:gnn_overview_edge}, Table~\ref{tab:gnn_overview_vertex} 
by observing the effects the hyper-parameters of each GNN model on training time per epoch and memory usage.

@@

Fig~\ref{fig:exp_absolute_training_time} compares the training time per epoch of each GNN model, the ranking is GaAN $\gg$ GAT $>$ GGNN $>$ GCN.
The time-consuming ranking is consistent with the complexity analysis. Because the number of edges in the graph generally far exceeds the number of nodes, 
the algorithm GAT with higher edge computing complexity is more time-consuming than the algorithm GGNN with high vertex computational complexity. At the same time, 
Fig~\ref{fig:exp_absolute_training_time} shows that the training time of individual epochs is abnormally high, which is mainly caused by profiling overhead and the GC pause of the Python interpreter.
This phenomenon confirms the necessity of removing abnormal epochs.

According to Table~\ref{tab:gnn_overview_edge}, Table~\ref{tab:gnn_overview_vertex}, the computational complexity of edge and vetex of each GNN model is linearly related to the 
hyper-parameters of each model(such as $d_{dim}$, $K$, etc at al). In order to prove this linear relationship, we measured the change of the training time of each GNN model with the hyperparameters.

@@

The computational complexity of GCN and GGNN is affected by the hidden vector dimension $d_{dim}$. $d_{dim}$ affects both the output hidden vector dimension fo Layer0 and
the input hidden vector dimension of Layer1(ie $d_{dim} = d^0_{out} = d^1_{in}$). Fig~\ref{fig:exp_hyperparameter_on_vertex_edge_phase_time_gcn}, Fig~\ref{fig:exp_hyperparameter_on_vertex_edge_phase_time_ggnn}
show that the training time of GCN and GGNN is affected by $d_{dim}$. As $d_{dim}$ increases, the training time increases linearly.

GAT uses a multi-head mechanism, and its computational complexity is affected by the input hidden vector dimension $d_{in}$, the hidden vector dimension $d_{head}$ of each head, and the number of heads $K$.
The output hidden vector dimension of each layer is $d_{out}=K d_{head}$.
Because in the GAT model $d^1_{in}=d^0_{out}$, adjusting $d_{head}$ and $K$ is equivalent to adjusting Layer1’s $d^1_{in}$.
Fig~\ref{fig:exp_hyperparameter_on_vertex_edge_phase_time_gat} shows that the GAT training time is affected by the hyperparameters $d_{head}$ and $K$, 
GAT training time increases linearly with $d_{head}$ and $K$.

GaAN also uses a multi-head mechanism, and its computational complexity is affected by $d_{in}$, $d_v$, $d_a$ and the number of heads $K$.
Fig~\ref{fig:exp_hyperparameter_on_vertex_edge_phase_time_gat} demonstrates that GaAN training time is affected by hyperparameters.
\\
The above experiments prove the complexity analysis results given in Table~\ref{tab:gnn_overview_edge}, Table~\ref{tab:gnn_overview_vertex}. 
And the training time of each GNN model algorithm increases linearly with the increase of hyperparameters.
When the hidden vector dimension $d_{in}$ is too low, the calculations involving hidden vectors account for a very low proportion of the total calculation time, resulting in an insignificant change in the total training time.
When the hidden vector dimension is large enough, the total training time increases linearly with $d_{in}$.

Fig~\ref{fig:exp_hyperparameter_memory_usage} also shows how each GNN model's GPU memory usage varies with the algorithm's hyperparameters.
With the increase of hyperparameters, GNN's memory usage also increases linearly.

@@

The experiment proves the effectiveness of the complexity analysis in Table~\ref{tab:gnn_overview_edge}, Table~\ref{tab:gnn_overview_vertex}. \textbf{GNN training time and memory usage are linearly related to hyperparameters}.
This allows algorithm engineers to use larger hyperparameters to increase the complexity of GNN without worrying about the time-consuming training and the explosive growth of memory usage.

@@

The goal of this experiment is to discover the computational performance bottleneck in GNN training
by decomposing the time-consuming training.

@@

For vertex calculation and edge calculation, Fig~\ref{fig:exp_vertex_edge_cal_proportion} shows the ratio of the vertex/edge calculation time of different GNN layers in each GNN
to the total training time (including forward, backward and evaluation stages). The GCN algorithm spends most of time on edge calculation in most datasets.
Only the \textit{cph} dataset is a special case, because the dimension of the input feature vector of this dataset is very high, resulting in the 
high vertex calculation time-consuming of Layer0. Because of its high vertex calculation complexity, GGNN makes its vertex calculation time-consuming\
ratio significantly higher than other algorithms, but in most datasets, edge calculation still occupies the majority of total time. 
Only in \textit{pub} and cam dataset, the edge calculation cost and the vertex calculation cost are close, because the average degree of the two datasets is low (only 4.5 and 2.8).
For the GAT and GaAN algorithms, because of their high edge calculation complexity, their edge calculation cost time is absolutely dominant. 
In summary, \textbf{edge calculation is the main time-consuming factor of GNN training}, especially in the case of more complicated edge calculation.

@@

Experiments also show that \textbf{the average degree of the dataset affects the time-consuming proportion of vertex/edge calculation}. 
Our fixed graph has 50k vertices, and the R-MAT generator is used to generate a random graph with an average degree between 10 and 100.
We measure the change of the time-consuming proportion of the vertex/edge calculation in each GNN model varies with the average degree of the graph,
as shown in Fig~\ref{fig:exp_avg_degree_on_vertex_edge_cal_time}. The time calculation of the edge increases linearly with the increase of the average degree,
\textbf{The time calculation of the edge calculation dominates the entire calculation time in most cases}. 
Only when the vertex calculation complexity is very high and the average degree is very low, the vertex calculation time can overtake the edge calculation time.
Therefore, \textbf{the focus of GNN training optimization should improve the efficiency of edge calculation}.

@@

\paragraph{Time-consuming decomposition analysis of edge calculation}
The edge calculation phase can be further decomposed into four steps of collect, message, aggregate and update,
as shown in Fig~\ref{fig:steps_in_edge_calculation}. The figure shows the edge calculation process of GNN layer $l$.
edge index is a matrix that holds the edge sets of the graph with shape $M*2$, where $M$ is the number of edges,
and the two columns of the matrix respectively store the source vertex and the target vertex of each edge.
The edge index remains unchanged throughout the calculation process.
The collect step is used to prepare the data structure required for edge calculation.
This step will use the vertex hidden vector of the GNN layer $\boldsymbol{h}_i^l (1 \leq i \leq N)$
as input, which is copied to the two layers of each edge according to the edge index, forming the input parameter tensor of the edge calculation function $\phi$
(including $\boldsymbol{h}_i^l$,$\boldsymbol{h}_j^ l$ and $\boldsymbol{e}_{j, i}^l$).
There is no calculation in this step, only data access is involved. The message step calls the function given by the user $\phi$ to
complete the edge calculation process, and get message vector of each edge $\boldsymbol{m}_{j, i}^l (\boldsymbol{e}_{j, i}^l \in E(G))$.
The aggregate step is based on the target of each edge, the message vector with the same target vertex is aggregated by the aggregation operator $\Sigma$,
and then each vertex aggregation vector $\boldsymbol{a}_i^l (1 \leq i \leq N)$ is obtained. 
The last update step is optional, and it can perform additional correction processing on the aggregated vector (for example, adding bias to GCN).
The aggregated vector $a_i^l$ after the update process will be input into the vertex calculation function $\gamma$ As an input parameter.

@@

We performed the execution time decomposition of the edge calculation process of each GNN algorithm on different datasets,
and the results are shown in Fig~\ref{fig:exp_edge_calc_decomposition}.
The time-consuming decomposition of the edge calculation of each GNN is relatively stable, 
and the dataset correlation is not strong. Although the collect step only performs data preparation,
it occupies a lot of execution time in all GNNs. For the message step, in the GNN (GAT and GaAN) with high edge calculation complexity, 
which occupies the absolute dominant. Although its edge calculation in GCN only has a simple multiplication operation, it still takes more than 20\% of the time; 
In GGNN, because its edge calculation function $\boldsymbol{m}_{j,i }^l=\boldsymbol{W}^l\boldsymbol{h}_{j}^l$ is only related to the source vertex, 
so in the implementation of PyG will be $\boldsymbol{W}^l\boldsymbol{h}_j The calculation of ^l$ is carried out before the edge calculation starts
(because this part of the calculation is only related to vertices, so we count the calculation into the vertex calculation phase).
The calculated result is cached, and it is directly read when the message step is performed. As a result, the message step of GGNN gets 0.
For the aggregate step, it occupies at least 35\% of training time of GNN (GCN and GGNN) with low edge computational complexity,
while the GNN with high edge computational complexity (GAT and GaAN), whose time consumption is close to the collect step,
and both are much lower than the message step. Experiments show that \textbf{for an algorithm with high edge calculation complexity,
the message step is the bottleneck of its performance, and it should be optimized}; and \textbf{For algorithms with low computational complexity,
optimizing the collect and aggregate steps can significantly reduce training time}.

@@

Fig~\ref{fig:exp_memory_usage_stage_amp} shows the peak memory usage of each stage when each GNN is trained on the cam dataset, the situation is similar on other datasets.
\textbf{The memory usage in GNN training is forward The peak in the phase and the backward phase}, 
because a large number of temporary calculation results will be generated in the forward phase,
and the key intermediate calculation results will be cached.
The cached intermediate results will be used in the gradient calculation of the backward phase.
Fig~\ref{fig:ggnn_vertex_func_computation_graph} shows the calculation graph of GGNN's vertex calculation function $\gamma$.
It can be seen that a large number of operators are involved in GGNN vertex calculation and a large number of intermediate calculation results are generated.The key calculation results will also be cached,
which is exacerbated Memory usage. Most of the peak memory usage in the loss phase comes from the intermediate calculation results of the cache.
With the end of the backward phase, the memory of the intermediate calculation results is released.
In the evaluation phase, there is no need to cache intermediate results for gradient calculations, Peak memory usage dropped significantly.

@@

It is worth noting that \textbf{the peak memory during GNN training far exceeds the memory usage of the dataset itself}.
The ratio of the highest peak memory usage during our training compared to the memory usage after data Load is
defined as the memory expansion ratio. Fig~\ref{fig:exp_memory_expansion_ratio} compares the memory expansion ratios
of different GNNs on different datasets. The expansion ratio of GCN is the lowest, between 5-14 times,
and the expansion ratio of GaAN is the highest, up to 101 times.
\textbf{Very high expansion ratio severely limits the data scalability of GNN, making GPU unable to process large-scale graph datasets},
especially restricting GNN with high edge calculation complexity.

@@

Fig~\ref{fig:exp_memory_expansion_ratio} also shows that the expansion ratio of the same GNN under the same hyperparameters changes with different datasets.
Because the input feature dimension of the \textit{cph} dataset is much higher than the hidden vector of the GNN layer The dimension of the input feature vector matrix
of the graph is much higher than the matrix size of the intermediate calculation result of the cache, 
so its expansion ratio is particularly low, and the cam dataset is the opposite. In order to measure the effect of the input feature vector dimension on the memory
expansion ratio , We randomly generated feature vectors of specific dimensions for different datasets.
Fig~\ref{fig:exp_memory_expension_ratio_input_feature_dimension} shows the change in the expansion ratio under different input feature vector dimensions.
\textbf{In the same GNN structure and hyperparameter settings, using a higher-dimensional input feature vector can reduce memory expansion ratio} .

@@

In the case of a fixed number of vertices in the graph, we use the R-MAT generator to generate random graphs with different average degrees (number of edges).
The Fig~\ref{fig:exp_memory_expansion_ratio_input_graph_number_of_edges} shows how the memory usage during training varies with the average degree.
\textbf{As the average degree increases, the peak memory usage increases linearly.
The intermediate results generated by the edge calculation gradually dominate, and the memory expansion ratio of each GNN gradually stabilizes}. 
The expansion ratio is affected by the complexity of the edge calculation.
Except for GGNN, the memory expansion ratio of other GNNs increases with the increase of degree. GGNN has a high vertex calculation complexity.
When the average degree is low, its memory expansion ratio is mainly affected by the intermediate results of vertex calculation;
when the average When the degree is increased, the memory expansion ratio is gradually determined by the edge calculation complexity.
Because GGNN has the lowest edge calculation complexity, its stable expansion ratio is the lowest.

@@

In the case of a fixed number of edges in the graph, we use the R-MAT generator to generate random graphs with different numbers of vertices
Fig~\ref{@fig:exp_memory_expansion_ratio_input_graph_number_of_vertices_fixed_edge} shows the change of memory usage with the number of vertices during training.
Except for GGNN, the number of vertices increases, the number of vertices increases, the number of vertices increases,
and the number of vertices increases. The size of the eigenvector matrix of the vertex input becomes larger, 
but the peak memory usage only increases slightly, which makes the proportion of memory expansion decrease to a certain extent. 
Only GGNN has a large number of intermediate calculation results during the vertex calculation process because of the high vertex calculation complexity. 
The memory expansion factor has a small increase. The experimental data shows that \textbf{the intermediate result produced by the edge calculation is the dominant factor
in memory usage, and the peak memory usage of each GNN increases linearly with the increase in the number of edges}.

@@

\paragraph{Bottleneck restricting data scalability}
\begin{itemize}
    \item \textbf{GPU memory capacity is the decisive factor limiting the scalability of the training dataset}.
    \item \textbf{GPU memory usage mainly comes from the intermediate calculation results generated during the calculation process,
     especially the intermediate calculation results of the edge calculation*. Because part of the intermediate calculation results
    will be cached to participate in backward calculation, high GPU memory usage runs through the forward and backward stages.}
    \item \textbf{GNN's peak memory usage during training can reach tens or even hundreds of times the size of the input data}. The limited memory capacity of the GPU severely limits the size of the input data that can be trained.
    \item \textbf{In the case of a fixed number of vertices, *The peak memory usage of GNN increases linearly with the increase of the number of edges of the graph*, *The proportion of memory expansion will gradually stabilize to a fixed value determined by the complexity of the edge calculation}.
    \item \textbf{In the case where the network structure and various hyperparameters are fixed, *using a higher-dimensional input feature vector can reduce the expansion ratio of GPU memory usage}.
\end{itemize}

@@

Fig~\ref{fig:exp_sampling_minibatch_graph_info} shows how the size of the subgraph sampled in the two sampling techniques varies with the batch size.
For Neighbor Sampler, the relative batch size is equal to the number of vertices sampled in the last layer of GNN relative to the top of the full graph the proportion of vertices.
For Cluster Sampler, the batch size is equal to the number of partitions sampled compared to the number of all partitions in the whole graph.
\textbf{Neighbor Sampler is very sensitive to the increase in batch size*. As the batch size increases, 
the sampling subgraph The number of vertices/edges and the average degree both increase rapidly and tend to stabilize}.
\textbf{Cluster Sampler is less sensitive to the increase in batch size*, the number of vertices/average degree increases
linearly with the increase in batch size, and the number of edges is in the batch When the size is relatively small,
it also increases linearly.}

It is worth noting that \textbf{the average degree of the subgraph sampled is much lower than the average degree of the whole graph,
especially when the relative batch size is low}. Taking Neighbor Sampler at a relative batch size of 6\% as an example,
The average degree of the \textit{amp} dataset is 31.1, but the average degree of the sampled subgraph is only 5.8,
which is much lower than the average degree of the full graph.
The average degree of the subgraph sampled by the Cluster Sampler is lower, only 3.0.
Fig~\ref{fig:exp_sampling_minibatch_degrees_distribution) shows the comparison of the distribution of the sampled subgraph with the original graph.
The vertex dregree distribution of subgraphs sampled by the two sampling techniques is lower than that of the original graph. 
The reason is that the degree distribution of vertex in the actual graph dataset follows a power-rate distribution and only a small number of vertices is very high,
which improves the average degree. In the sampling process, because the sampling method will limit the neighborhood size of the vertex, thereby reducing the upper limit of the vertex degree,
so that the average degree drops significantly. Combined with the experimental results in section 4.2, the decrease in the average degree of sampled subgraph will increase the time-consuming
proportion of vertex calculation. For GGNN with high vertex calculation complexity, vertex calculation will replace edge calculation and become a performance bottleneck.

@@

Fig~\ref{fig:exp_sampling_batch_train_time} shows the change of the training time of each batch on the \textit{amc} and \textit{fli} datasets with the batch size after using the sampling technique.
For the Neighbor Sampler, the time spent in the training phase is significantly reduced compared to full-batch training only when the batch size is very small. 
When the batch size is particularly small, the training time is significantly reduced. However, because the sampling itself has additional overhead, 
and there is also an overhead in transferring the sampled subgraph to the GPU, the training time of the entire batch may exceed the full graph. 
training is time-consuming. For Cluster Sampler, the sampled subgraph is smaller than Neighbor Sampler under the same relative batch size,
which makes the effect of sampling technology more obvious. But when the relative batch size increases, 
the increment by additional cost of sampling technology is very obvious. When the relative batch size $\geq$ 25\%,
the total time of each batch of the sampling technique even exceeds the time of the full graph training. 
Experiments show that \textbf{the implementation of sampling technology in PyG at this stage is very inefficient and the extra overhead is high}.
When the batch size is slightly larger, the additional cost of sampling technology accounts for more than 50\% of the total training cost of the entire batch.
\textbf{Sampling technology can reduce training time only on very small batches}.

@@

However, \textbf{the advantage of the sampling technique is that it can greatly reduce the peak memory usage overhead,
making large-scale graph neural network training possible}. 
Fig~\ref{fig:exp_sampling_memory_usage} shows how the peak memory usage changes with the relative batch size during the training process after sampling technology is used.
After sampling sampling technology, peak memory usage is greatly reduced and the decrease of Cluster Sampler is greater than that of Neighbor Sampler.

>>>>

To verify the accuracy of complexity analysis in \tablename~\ref{tab:gnn_overview_edge}, \tablename~\ref{tab:gnn_overview_vertex}, we 
design following experiments to observe the effects the hyper-parameters on training time and memory usage of a GNN. Especially, the memory usage
points to GPU memory usage, same to below.

\paragraph{the neccessity of removing abnormal epochs}
The absolute training time of four GNN models is shown in \figurename~\ref{fig:exp_absolute_training_time}.
In figure, we can conclude that the time ranking is GaAN $\gg$ GAT $>$ GGNN $>$ GCN. In all datasets, the number of edges is greater than that of nodes, so the correlation of training time between GAT and GGNN is
GAT $>$ GGNN. And the remaining relationship is consistent with complexity analysis. In addition, we find that there are some outliers in \figurename~\ref{fig:exp_absolute_training_time} because of profiling overhead
and GC pause of the Python interpreter. So in all the experiments below, we will remove abnormal epochs and use the average of remaining epochs to analysis.

@@

\paragraph{the effects of hyper-parameters on training time}

According to \tablename~\ref{tab:gnn_overview_edge}, \tablename~\ref{tab:gnn_overview_vertex}, the computational complexity of edge and vetex of each GNN is linearly related to the 
hyper-parameters of each model(such as $d_{dim}$, $K$, etc at al). 
In order to prove this linear relationship, we measured the change of the training time of each GNN model with the hyperparameters.

@@

For GCN and GGNN, the computational complexity is affected by the hidden vector dimension $d_{dim}$. $d_{dim}$ affects both the output hidden vector dimension of Layer0 and
the input hidden vector dimension of Layer1($d_{dim} = d^0_{out} = d^1_{in}$).
\figurename~\ref{fig:exp_hyperparameter_on_vertex_edge_phase_time_gcn} and \figurename~\ref{fig:exp_hyperparameter_on_vertex_edge_phase_time_ggnn}
display that the training time of GCN and GGNN is affected by $d_{dim}$. As $d_{dim}$ increases, the training time increases linearly.

For GAT,  which uses a multi-head mechanism, its computational complexity is affected by the input hidden vector dimension $d_{in}$ and the hidden vector dimension $d_{head}$ of each head, and the number of heads $K$.
The output hidden vector dimension of each layer is $d_{out}=K d_{head}$.
Because of GAT model structure $d^1_{in}=d^0_{out}$, adjusting $d_{head}$ and $K$ is equivalent to adjusting Layer1’s $d^1_{in}$.
\figurename~\ref{fig:exp_hyperparameter_on_vertex_edge_phase_time_gat} shows that the GAT training time increases linearly with $d_{head}$ and $K$, two.

For GaAN, which also has a multi-head mechanism, its computational complexity is affected by $d_{in}$, $d_v$, $d_a$ and the number of heads $K$.
\figurename~\ref{fig:exp_hyperparameter_on_vertex_edge_phase_time_gat} demonstrates that GaAN training time is linearly affected by hyperparameters.
\\
The above experiments prove the complexity analysis results given in \tablename~\ref{tab:gnn_overview_edge}, \tablename~\ref{tab:gnn_overview_vertex}. 
The training time of each GNN model increases linearly with the increase of hyperparameters.
When the hidden vector dimension $d_{in}$ is too low, the calculations involving hidden vectors account for a very low proportion of the total calculation time,
resulting in an insignificant change in the total training time.
When the hidden vector dimension is large enough, the total training time increases linearly with $d_{in}$.

\paragraph{the effects of hyper-parameters on memory usage}

\figurename~\ref{fig:exp_hyperparameter_memory_usage} shows how each GNN model's memory usage varies with the algorithm's hyperparameters.
With the increase of hyperparameters, GNN's memory usage also increases linearly.

@@

\paragraph{Summary}

We can answer the question Q1 \ref{sec:experimental_env}: the complexity analysis in \tablename~\ref{tab:gnn_overview_edge}, \tablename~\ref{tab:gnn_overview_vertex} is valid. 
\textbf{GNN training time and memory usage are linearly related to hyperparameters}.
This allows algorithm engineers to use larger hyperparameters to increase the complexity of GNN without worrying about the time-consuming training and the explosive growth of memory usage.

@@

In thi subsection, we aims to find what stage is the most time-consuming stage in GNN training by decomposing the time-consuming training.

@@

For vertex calculation and edge calculation, \figurename~\ref{fig:exp_vertex_edge_cal_proportion} shows the ratio of the vertex/edge calculation time of different GNN layers in each GNN
to the total training time (including forward, backward and evaluation stages). 
The GCN algorithm spends most of time on edge calculation in most datasets except \textit{cph} dataset on account of
the dimension of the input feature vector of this dataset is very high, resulting in the high vertex calculation time-consuming of Layer0.
Because of its high vertex calculation complexity, GGNN makes its vertex calculation time-consuming\
ratio significantly higher than other algorithms, but in most datasets, edge calculation still occupies the majority of total time. 
Only in \textit{pub} and cam dataset, the edge calculation cost and the vertex calculation cost are close for GGNN 
because the average degree of the two datasets is low (only 4.5 and 2.8).
For the GAT and GaAN algorithms, because of their high edge calculation complexity, their edge calculation cost time is absolutely dominant. 
In summary, \textbf{edge calculation is the main time-consuming factor of GNN training}, especially in the case of more complicated edge calculation.

@@

Experiments also show that \textbf{the average degree of the dataset affects the time-consuming proportion of vertex/edge calculation}. 
We use the R-MAT generator to generate random graphs with an average degree between 10 and 100. The number of nodes of these graphs is fixed at 50k.
Then we measure the change of the time-consuming proportion of the vertex/edge calculation in each GNN model varies with the average degree of the graph,
as shown in \figurename~\ref{fig:exp_avg_degree_on_vertex_edge_cal_time}. The edge calculation time increases linearly with the increase of the average degree,
\textbf{The edge calculation time dominates the entire calculation time in most cases}. 
Only when the vertex calculation complexity is very high and the average degree is very low, the vertex calculation time can overtake the edge calculation time.
Therefore, \textbf{the focus of GNN training optimization should improve the efficiency of edge calculation}.

@@

\paragraph{Time-consuming decomposition analysis of edge calculation}

The edge calculation phase can be further decomposed into four steps of collect, message, aggregate and update,
as shown in \figurename~\ref{fig:steps_in_edge_calculation}. The figure shows the edge calculation process of GNN layer $l$.
edge index is a matrix that holds the edge sets of the graph with shape $M*2$, where $M$ is the number of edges,
and the two columns of the matrix respectively store the source vertex and the target vertex of each edge.
The edge index remains unchanged throughout the calculation process.
The collect step is used to prepare the data structure required for edge calculation and will use the vertex hidden vector of the GNN layer $\boldsymbol{h}_i^l (1 \leq i \leq N)$
as input, which is copied to the two layers of each edge according to the edge index, forming the input parameter tensor of the edge calculation function $\phi$
(including $\boldsymbol{h}_i^l$,$\boldsymbol{h}_j^ l$ and $\boldsymbol{e}_{j, i}^l$).
There is no calculation in this step, only data access is involved. 
The message step calls the function given by the user $\phi$ to complete the edge calculation process, and get message vector of each edge $\boldsymbol{m}_{j, i}^l (\boldsymbol{e}_{j, i}^l \in E(G))$.
The aggregate step is based on the target of each edge, the message vector with the same target vertex is aggregated by the aggregation operator $\Sigma$,
and then each vertex aggregation vector $\boldsymbol{a}_i^l (1 \leq i \leq N)$ is obtained. 
The last update step is optional, and it can perform additional correction processing on the aggregated vector (for example, adding bias to GCN).
The aggregated vector $a_i^l$ after the update process will be input into the vertex calculation function $\gamma$ as an input parameter.

@@

We performed the execution time decomposition of the edge calculation process of each GNN algorithm on different datasets,
and the results are shown in \figurename~\ref{fig:exp_edge_calc_decomposition}.
The time-consuming decomposition of the edge calculation of each GNN is relatively stable and rarely affected by dataset.
Although the collect step only performs data preparation, it occupies a lot of execution time in all GNNs.
The message step in the GNN (GAT and GaAN) with high edge calculation complexity occupies the absolute dominant.
Although its edge calculation in GCN only has a simple multiplication operation, it still takes more than 20\% of the time; 
In GGNN, because its edge calculation function $\boldsymbol{m}_{j,i }^l=\boldsymbol{W}^l\boldsymbol{h}_{j}^l$ is only related to the source vertex, 
so in the implementation of PyG will be $\boldsymbol{W}^l\boldsymbol{h}_j The calculation of ^l$ is carried out before the edge calculation starts
(because this part of the calculation is only related to vertices, so we count the calculation into the vertex calculation phase).
The calculated result is cached, and it is directly read when the message step is performed. As a result, the message step of GGNN gets 0.
For the aggregate step, it occupies at least 35\% of training time of GNN (GCN and GGNN) with low edge computational complexity,
while the GNN with high edge computational complexity (GAT and GaAN), whose time consumption is close to the collect step,
and both are much lower than the message step. Experiments show that \textbf{for an algorithm with high edge calculation complexity,
the message step is the bottleneck of its performance, and it should be optimized}; and \textbf{For algorithms with low computational complexity,
optimizing the collect and aggregate steps can significantly reduce training time}.

@@

\figurename~\ref{fig:exp_memory_usage_stage_amp} shows the peak memory usage of each stage when each GNN is trained on the cam dataset, the situation is similar on other datasets.
\textbf{The memory usage in GNN training is forward The peak in the phase and the backward phase}, 
because a large number of temporary calculation results will be generated in the forward phase,
and the key intermediate calculation results will be cached.
The cached intermediate results will be used in the gradient calculation of the backward phase.
\figurename~\ref{fig:ggnn_vertex_func_computation_graph} shows the calculation graph of GGNN's vertex calculation function $\gamma$.
It can be seen that a large number of operators are involved in GGNN vertex calculation and a large number of intermediate calculation results are generated.The key calculation results will also be cached,
which is exacerbated Memory usage. Most of the peak memory usage in the loss phase comes from the intermediate calculation results of the cache.
With the end of the backward phase, the memory of the intermediate calculation results is released.
In the evaluation phase, there is no need to cache intermediate results for gradient calculations, Peak memory usage dropped significantly.

@@

It is worth noting that \textbf{the peak memory during GNN training far exceeds the memory usage of the dataset itself}.
The ratio of the highest peak memory usage during our training compared to the memory usage after data Load is
defined as the memory expansion ratio. \figurename~\ref{fig:exp_memory_expansion_ratio} compares the memory expansion ratios
of different GNNs on different datasets. The expansion ratio of GCN is the lowest, between 5-14 times,
and the expansion ratio of GaAN is the highest, up to 101 times.
\textbf{Very high expansion ratio severely limits the data scalability of GNN, making GPU unable to process large-scale graph datasets},
especially restricting GNN with high edge calculation complexity.

@@

\figurename~\ref{fig:exp_memory_expansion_ratio} also shows that the expansion ratio of the same GNN under the same hyperparameters changes with different datasets.
Because the input feature dimension of the \textit{cph} dataset is much higher than the hidden vector of the GNN layer The dimension of the input feature vector matrix
of the graph is much higher than the matrix size of the intermediate calculation result of the cache, 
so its expansion ratio is particularly low, and the cam dataset is the opposite. In order to measure the effect of the input feature vector dimension on the memory
expansion ratio , We randomly generated feature vectors of specific dimensions for different datasets.
\figurename~\ref{fig:exp_memory_expension_ratio_input_feature_dimension} shows the change in the expansion ratio under different input feature vector dimensions.
\textbf{In the same GNN structure and hyperparameter settings, using a higher-dimensional input feature vector can reduce memory expansion ratio} .

@@

In the case of a fixed number of vertices in the graph, we use the R-MAT generator to generate random graphs with different average degrees (number of edges).
The \figurename~\ref{fig:exp_memory_expansion_ratio_input_graph_number_of_edges} shows how the memory usage during training varies with the average degree.
\textbf{As the average degree increases, the peak memory usage increases linearly.
The intermediate results generated by the edge calculation gradually dominate, and the memory expansion ratio of each GNN gradually stabilizes}. 
The expansion ratio is affected by the complexity of the edge calculation.
Except for GGNN, the memory expansion ratio of other GNNs increases with the increase of degree. GGNN has a high vertex calculation complexity.
When the average degree is low, its memory expansion ratio is mainly affected by the intermediate results of vertex calculation;
when the average When the degree is increased, the memory expansion ratio is gradually determined by the edge calculation complexity.
Because GGNN has the lowest edge calculation complexity, its stable expansion ratio is the lowest.

@@

In the case of a fixed number of edges in the graph, we use the R-MAT generator to generate random graphs with different numbers of vertices
\figurename~\ref{@fig:exp_memory_expansion_ratio_input_graph_number_of_vertices_fixed_edge} shows the change of memory usage with the number of vertices during training.
Except for GGNN, the number of vertices increases, the number of vertices increases, the number of vertices increases,
and the number of vertices increases. The size of the eigenvector matrix of the vertex input becomes larger, 
but the peak memory usage only increases slightly, which makes the proportion of memory expansion decrease to a certain extent. 
Only GGNN has a large number of intermediate calculation results during the vertex calculation process because of the high vertex calculation complexity. 
The memory expansion factor has a small increase. The experimental data shows that \textbf{the intermediate result produced by the edge calculation is the dominant factor
in memory usage, and the peak memory usage of each GNN increases linearly with the increase in the number of edges}.

@@

\paragraph{Bottleneck restricting data scalability}
\begin{itemize}
    \item \textbf{GPU memory capacity is the decisive factor limiting the scalability of the training dataset}.
    \item \textbf{GPU memory usage mainly comes from the intermediate calculation results generated during the calculation process,
     especially the intermediate calculation results of the edge calculation. Because part of the intermediate calculation results
    will be cached to participate in backward calculation, high GPU memory usage runs through the forward and backward stages.}
    \item \textbf{GNN's peak memory usage during training can reach tens or even hundreds of times the size of the input data. The limited memory capacity of the GPU severely limits the size of the input data that can be trained}.
    \item \textbf{In the case of a fixed number of vertices, The peak memory usage of GNN increases linearly with the increase of the number of edges of the graph, The proportion of memory expansion will gradually stabilize to a fixed value determined by the complexity of the edge calculation}.
    \item \textbf{In the case where the network structure and various hyperparameters are fixed, using a higher-dimensional input feature vector can reduce the expansion ratio of GPU memory usage}.
\end{itemize}

@@

\figurename~\ref{fig:exp_sampling_minibatch_graph_info} shows how the size of the subgraph sampled in the two sampling techniques varies with the batch size.
For Neighbor Sampler, the relative batch size is equal to the number of vertices sampled in the last layer of GNN relative to the top of the full graph the proportion of vertices.
For Cluster Sampler, the batch size is equal to the number of partitions sampled compared to the number of all partitions in the whole graph.
\textbf{Neighbor Sampler is very sensitive to the increase in batch size*. As the batch size increases, 
the sampling subgraph The number of vertices/edges and the average degree both increase rapidly and tend to stabilize}.
\textbf{Cluster Sampler is less sensitive to the increase in batch size*, the number of vertices/average degree increases
linearly with the increase in batch size, and the number of edges is in the batch When the size is relatively small,
it also increases linearly.}

It is worth noting that \textbf{the average degree of the subgraph sampled is much lower than the average degree of the whole graph,
especially when the relative batch size is low}. Taking Neighbor Sampler at a relative batch size of 6\% as an example,
The average degree of the \textit{amp} dataset is 31.1, but the average degree of the sampled subgraph is only 5.8,
which is much lower than the average degree of the full graph.
The average degree of the subgraph sampled by the Cluster Sampler is lower, only 3.0.
\figurename~\ref{fig:exp_sampling_minibatch_degrees_distribution} shows the comparison of the distribution of the sampled subgraph with the original graph.
The vertex dregree distribution of subgraphs sampled by the two sampling techniques is lower than that of the original graph. 
The reason is that the degree distribution of vertex in the actual graph dataset follows a power-rate distribution and only a small number of vertices is very high,
which improves the average degree. In the sampling process, because the sampling method will limit the neighborhood size of the vertex, thereby reducing the upper limit of the vertex degree,
so that the average degree drops significantly. Combined with the experimental results in section 4.2, the decrease in the average degree of sampled subgraph will increase the time-consuming
proportion of vertex calculation. For GGNN with high vertex calculation complexity, vertex calculation will replace edge calculation and become a performance bottleneck.

@@

\figurename~\ref{fig:exp_sampling_batch_train_time} shows the change of the training time of each batch on the \textit{amc} and \textit{fli} datasets with the batch size after using the sampling technique.
For the Neighbor Sampler, the time spent in the training phase is significantly reduced compared to full-batch training only when the batch size is very small. 
When the batch size is particularly small, the training time is significantly reduced. However, because the sampling itself has additional overhead, 
and there is also an overhead in transferring the sampled subgraph to the GPU, the training time of the entire batch may exceed the full graph. 
training is time-consuming. For Cluster Sampler, the sampled subgraph is smaller than Neighbor Sampler under the same relative batch size,
which makes the effect of sampling technology more obvious. But when the relative batch size increases, 
the increment by additional cost of sampling technology is very obvious. When the relative batch size $\geq$ 25\%,
the total time of each batch of the sampling technique even exceeds the time of the full graph training. 
Experiments show that \textbf{the implementation of sampling technology in PyG at this stage is very inefficient and the extra overhead is high}.
When the batch size is slightly larger, the additional cost of sampling technology accounts for more than 50\% of the total training cost of the entire batch.
\textbf{Sampling technology can reduce training time only on very small batches}.

@@

However, \textbf{the advantage of the sampling technique is that it can greatly reduce the peak memory usage overhead,
making large-scale graph neural network training possible}. 
\figurename~\ref{fig:exp_sampling_memory_usage} shows how the peak memory usage changes with the relative batch size during the training process after sampling technology is used.
After sampling sampling technology, peak memory usage is greatly reduced and the decrease of Cluster Sampler is greater than that of Neighbor Sampler.
